Expt 5: Mix of PD and BoS with spikes and two types of PD trials

### 2025-06-14 code/analysis_oct2024.Rmd
do models for ms

>>>>

# Subgroup analysis

As an exploratory thing, what if we look at the people who talked *a lot* or *a moderate amount*

high = 40%+ of trials (10 games)

med = 10% - 39% of trials (10 games)

minimal = < 10% of trials (33 games)

no = couldn't chat (53 games)

```{r}

chat_subgroup <- how_many_chat |> mutate(chat_subgroup=case_when(
  pct_chat>.4 ~ "high",
  pct_chat>.09 ~ "med",
  T ~ "min"
))

subgroup_rounds <- rounds |> left_join(chat_subgroup) |>
  mutate(chat_subgroup=ifelse(is.na(chat_subgroup), "no", chat_subgroup))

subgroup_game_chat <- game_chat |> left_join(chat_subgroup) |>
  mutate(chat_subgroup=ifelse(is.na(chat_subgroup), "no", chat_subgroup))
```

```{r}
bonuses <- subgroup_rounds %>%
  group_by(playerId, gameId, chat_subgroup, gametype) %>%
  summarize(payoff = mean(payoff))

ggplot(bonuses, aes(x = chat_subgroup, y = payoff, color = str_c(chat_subgroup))) +
  geom_jitter(width = .1, alpha = .1) +
  facet_grid(. ~ gametype) +
  scale_color_viridis(discrete = T) +
  labs(y = "mean_reward", x = "chat usage") +
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black", geom="errorbar") +
    stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width = .3), size = .7, color = "black") +
  theme(legend.position = "none") +
  coord_cartesian(ylim = c(0, 20)) +
  labs(y = "Mean reward / trial") +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14)
  )

```{r}
ggplot(subgroup_game_chat, aes(x = str_c(chat_subgroup, "_", is_chat), y = mean_payout)) +
    geom_jitter(aes(color=chat_subgroup), alpha=.2)+

  stat_summary(fun.data = "mean_cl_boot", geom="errorbar") +
    stat_summary(fun.data = "mean_cl_boot") +

facet_wrap(~gametype) +
  geom_smooth(method = "lm")+
  theme(legend.position = "none")
```

## Option

```{r}
subgroup_outcome <- subgroup_rounds %>%
  select(gameId, repNum, response, chat_subgroup, gametype, role) |>
  pivot_wider(names_from = role, values_from = response) %>%
  mutate(outcome = str_c(p1, p2)) %>%
  left_join(game_chat)
```

### BoS
```{r}
subgroup_outcome %>%
  filter(gametype == "BoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_subgroup, "_", is_chat), y = outcome_parity, color=chat_subgroup)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")
```

### Spike BoS

```{r}
subgroup_outcome %>%
  filter(gametype == "spikeBoS") %>%
  mutate(outcome_parity = ifelse(outcome %in% c("AA", "BB"), 1, 0)) %>%
  ggplot(aes(x = str_c(chat_subgroup, "_", is_chat), y = outcome_parity, color=chat_subgroup)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) +
  labs(title = "Spiked BoS:fraction choosing good outcomes (AA or BB)", y = "Good outcome?")

### easy PD

```{r}
subgroup_outcome %>%
  filter(gametype == "easyPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = str_c(chat_subgroup, "_", is_chat), y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) 
```

### hard PD

```{r}
subgroup_outcome %>%
  filter(gametype == "hardPD") %>%
  mutate(AA = ifelse(outcome %in% c("AA"), 1, 0), 
         AB.BA =ifelse(outcome %in% c("AB", "BA"), 1, 0), 
         BB = ifelse(outcome %in% c("BB"), 1, 0)) %>%
  pivot_longer(AA:BB, names_to="type", values_to="selected") |> 
  ggplot(aes(x = str_c(chat_subgroup, "_", is_chat), y = selected, color=type)) +
  stat_summary(fun.data = "mean_cl_boot") +
  geom_hline(aes(yintercept = .5)) 
```
So if you do use the chat, you tend to get the best option. (Using chat means that easy and hard PD look different)
